\Large Taking uncertainty seriously \newline \normalsize A Bayesian approach to word embedding bias estimation

### 2022-03-06 presentations/presentationBoston/presentationBoston.Rmd
Adds weat simulations

### 50k simulations (same parameters)

\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
set.seed(123)
biasesNull <- numeric(50000)
for(i in 1:50000){
t1 <- data.frame(A  = rnorm(5,0,0.05), B = rnorm(5,0,0.05))
t2 <- data.frame(A  = rnorm(5,0,0.05), B = rnorm(5,0,0.05))
t3 <- data.frame(A  = rnorm(5,0,0.05), B = rnorm(5,0,0.05))
t4 <- data.frame(A  = rnorm(5,0,0.05), B = rnorm(5,0,0.05))

factor <- sd(c(s(t1),s(t2),s(t3),s(t4)))
numerator <-  mean(s(t1),s(t2)) - mean(s(t3),s(t4))
biasesNull[i]  <- numerator / factor
}

ggplot()+geom_histogram(aes(x=biasesNull, y = ..density..), alpha = 0.6, bins=50)+
  theme_tufte(base_size = 15)+labs(title="50k biases for identical means and sd =.05")+ xlab("bias")
```
\normalsize

>>>>

## The problem with pre-averaging on real set-up

### One simulation on real set-up (no similarities and small std)

\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "70%"}
set.seed(123456)
t1 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t2 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t3 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t4 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t5 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t6 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t7 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t8 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t9 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t10 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t11 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t12 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t13 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t14 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t15 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t16 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))

xlimits <- c(-.3,.3)

t8Plot <- ggplot(t8)+geom_point(aes(y = 0, x = A), size = 1, alpha = 0.5)+geom_point(aes(y = 0.1, x = B), col = "skyblue", size = 1, alpha = 0.5)+theme_tufte()+xlab("similarity")+scale_y_continuous(limits = c(-.05,.15),breaks = c(0,0.1), labels = c("A","B"))+ylab("group") + geom_vline(aes(xintercept = mean(A)), size = 0.3) + geom_vline(aes(xintercept = mean(B)), size = 0.3, col = "skyblue") +ggtitle("Similarities for term t3 in Y")+xlim(xlimits)

t9Plot <- ggplot(t9)+geom_point(aes(y = 0, x = A), size = 1, alpha = 0.5)+geom_point(aes(y = 0.1, x = B), col = "skyblue", size = 1, alpha = 0.5)+theme_tufte()+xlab("similarity")+scale_y_continuous(limits = c(-.05,.15),breaks = c(0,0.1), labels = c("A","B"))+ylab("group") + geom_vline(aes(xintercept = mean(A)), size = 0.3) + geom_vline(aes(xintercept = mean(B)), size = 0.3, col = "skyblue") +ggtitle("Similarities for term t4 in Y")+xlim(xlimits)

grid.arrange(t1Plot,t2Plot, t8Plot, t9Plot, ncol=2)
```

## The problem with pre-averaging on real set-up

### 10k simulations (same parameters)

\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}

set.seed(123)
biasesD0 <- numeric(10000)

for(i in 1:10000){
  t1 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t2 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t3 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t4 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t5 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t6 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t7 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t8 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t9 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t10 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t11 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t12 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t13 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t14 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t15 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t16 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))

  factorD0 <- sd(c(s(t1),s(t2),s(t3),s(t4),s(t5),s(t6),s(t7),s(t8),s(t9),s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16)))
  numeratorD0 <-  mean(s(t1),s(t2),s(t3),s(t4),s(t5),s(t6),s(t7),s(t8)) - mean(s(t9),s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16))
  biasesD0[i] <- numeratorD0/factorD0
}

ggplot()+geom_histogram(aes(x=biasesD0, y = ..density..), alpha = 0.6, bins=50)+
  theme_tufte()+labs(title="10k biases for same means and sd =.08")+ xlab("bias")

```
\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
quantile <- quantile(biasesD0, c(0.025, 0.975))
absOver182 <- mean(abs(biasesD0)>=1.81)
```
\normalsize
\pause

- The 95% of the data is in range `r round(quantile, 3)`
- `r round(absOver182*100, 2)`% of the absolute values are above 1.81

## The problem with pre-averaging on real set-up

### 10k simulations with similarity equal to 0.1

\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
set.seed(123456)
biasesD1 <- numeric(10000)

for(i in 1:10000){
  t1 <- data.frame(A  = rnorm(8,.1,0.08), B = rnorm(8,0,0.08))
  t2 <- data.frame(A  = rnorm(8,.1,0.08), B = rnorm(8,0,0.08))
  t3 <- data.frame(A  = rnorm(8,.1,0.08), B = rnorm(8,0,0.08))
  t4 <- data.frame(A  = rnorm(8,.1,0.08), B = rnorm(8,0,0.08))
  t5 <- data.frame(A  = rnorm(8,.1,0.08), B = rnorm(8,0,0.08))
  t6 <- data.frame(A  = rnorm(8,.1,0.08), B = rnorm(8,0,0.08))
  t7 <- data.frame(A  = rnorm(8,.1,0.08), B = rnorm(8,0,0.08))
  t8 <- data.frame(A  = rnorm(8,.1,0.08), B = rnorm(8,0,0.08))
  t9 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.1,0.08))
  t10 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.1,0.08))
  t11 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.1,0.08))
  t12 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.1,0.08))
  t13 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.1,0.08))
  t14 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.1,0.08))
  t15 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.1,0.08))
  t16 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.1,0.08))

  factorD1 <- sd(c(s(t1),s(t2),s(t3),s(t4),s(t5),s(t6),s(t7),s(t8),s(t9),s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16)))
  numeratorD1 <-  mean(s(t1),s(t2),s(t3),s(t4),s(t5),s(t6),s(t7),s(t8)) - mean(s(t9),s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16))
  biasesD1[i] <- numeratorD1/factorD1
}

ggplot()+geom_histogram(aes(x=biasesD1, y = ..density..), alpha = 0.6, bins=50)+
  theme_tufte()+labs(title="10k biases for different means 0.1 and sd =.08")+ xlab("bias")

```
\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
quantiles1 <- quantile(biasesD1, c(0.025, 0.975))
absMean1 <- mean(abs(biasesD1)>=1.81)
```
\normalsize
\pause

- The 95% of the data is in range `r round(quantiles1, 3)`
- `r round(absMean1*100, 2)`% of the absolute values are above 1.81

## The problem with pre-averaging on real set-up

### 10k simulations with similarity equal to 0.4

\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
set.seed(123456)
biasesD4 <- numeric(10000)

for(i in 1:10000){
  t1 <- data.frame(A  = rnorm(8,.4,0.08), B = rnorm(8,0,0.08))
  t2 <- data.frame(A  = rnorm(8,.4,0.08), B = rnorm(8,0,0.08))
  t3 <- data.frame(A  = rnorm(8,.4,0.08), B = rnorm(8,0,0.08))
  t4 <- data.frame(A  = rnorm(8,.4,0.08), B = rnorm(8,0,0.08))
  t5 <- data.frame(A  = rnorm(8,.4,0.08), B = rnorm(8,0,0.08))
  t6 <- data.frame(A  = rnorm(8,.4,0.08), B = rnorm(8,0,0.08))
  t7 <- data.frame(A  = rnorm(8,.4,0.08), B = rnorm(8,0,0.08))
  t8 <- data.frame(A  = rnorm(8,.4,0.08), B = rnorm(8,0,0.08))
  t9 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.4,0.08))
  t10 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.4,0.08))
  t11 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.4,0.08))
  t12 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.4,0.08))
  t13 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.4,0.08))
  t14 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.4,0.08))
  t15 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.4,0.08))
  t16 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.4,0.08))

  factorD4 <- sd(c(s(t1),s(t2),s(t3),s(t4),s(t5),s(t6),s(t7),s(t8),s(t9),s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16)))
  numeratorD4 <-  mean(s(t1),s(t2),s(t3),s(t4),s(t5),s(t6),s(t7),s(t8)) - mean(s(t9),s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16))
  biasesD4[i] <- numeratorD4/factorD4
}

ggplot()+geom_histogram(aes(x=biasesD4, y = ..density..), alpha = 0.6, bins=50)+
  theme_tufte()+labs(title="10k biases for different means 0.4 and sd =.08")+ xlab("bias")

```
\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
quantiles4 <- quantile(biasesD4, c(0.025, 0.975))
absMean4 <- mean(abs(biasesD4)>=1.81)
```
\normalsize
\pause

- The 95% of the data is in range `r round(quantiles4, 3)`
- `r round(absMean4*100, 2)`% of the absolute values are above 1.81

### 2022-03-06 presentations/presentationBoston/presentationBoston.tex
Adds weat simulations

\begin{frame}{The problem with pre-averaging}
\protect\hypertarget{the-problem-with-pre-averaging-2}{}
\begin{block}{50k simulations (same parameters)}
\protect\hypertarget{k-simulations-same-parameters}{}
\vspace{1mm}
\footnotesize

\begin{center}\includegraphics[width=0.8\linewidth]{presentationBoston_files/figure-beamer/unnamed-chunk-8-1} \end{center}
\normalsize
\end{block}
\end{frame}

>>>>

\begin{frame}{The problem with pre-averaging on real set-up}
\protect\hypertarget{the-problem-with-pre-averaging-on-real-set-up}{}
\begin{block}{One simulation on real set-up (no similarities and small
std)}
\protect\hypertarget{one-simulation-on-real-set-up-no-similarities-and-small-std}{}
\vspace{1mm}
\footnotesize

\begin{center}\includegraphics[width=0.7\linewidth]{presentationBoston_files/figure-beamer/unnamed-chunk-8-1} \end{center}
\end{block}
\end{frame}

\begin{frame}{The problem with pre-averaging on real set-up}
\protect\hypertarget{the-problem-with-pre-averaging-on-real-set-up-1}{}
\begin{block}{10k simulations (same parameters)}
\protect\hypertarget{k-simulations-same-parameters}{}
\vspace{1mm}
\footnotesize

\begin{center}\includegraphics[width=0.8\linewidth]{presentationBoston_files/figure-beamer/unnamed-chunk-9-1} \end{center}
\vspace{1mm}
\footnotesize

\begin{itemize}
\tightlist
\item
  The 95\% of the data is in range -2.763, 2.698
\item
  21.38\% of the absolute values are above 1.81
\end{itemize}
\end{block}
\end{frame}

\begin{frame}{The problem with pre-averaging on real set-up}
\protect\hypertarget{the-problem-with-pre-averaging-on-real-set-up-2}{}
\begin{block}{10k simulations with similarity equal to 0.1}
\protect\hypertarget{k-simulations-with-similarity-equal-to-0.1}{}
\vspace{1mm}
\footnotesize

\begin{center}\includegraphics[width=0.8\linewidth]{presentationBoston_files/figure-beamer/unnamed-chunk-11-1} \end{center}
\vspace{1mm}
\footnotesize

\begin{itemize}
\tightlist
\item
  The 95\% of the data is in range 0.851, 2.764
\item
  51.3\% of the absolute values are above 1.81
\end{itemize}
\end{block}
\end{frame}

\begin{frame}{The problem with pre-averaging on real set-up}
\protect\hypertarget{the-problem-with-pre-averaging-on-real-set-up-3}{}
\begin{block}{10k simulations with similarity equal to 0.4}
\protect\hypertarget{k-simulations-with-similarity-equal-to-0.4}{}
\vspace{1mm}
\footnotesize

\begin{center}\includegraphics[width=0.8\linewidth]{presentationBoston_files/figure-beamer/unnamed-chunk-13-1} \end{center}
\vspace{1mm}
\footnotesize

\begin{itemize}
\tightlist
\item
  The 95\% of the data is in range 1.679, 2.185
\item
  82.9\% of the absolute values are above 1.81
\end{itemize}
\end{block}
\end{frame}
